\vspace{-1cm} \bf Approximate DP and Regret Analysis \vspace{-1.3cm}

### 2018-08-07 Draft.tex
Sigmetrics format

\section{Probing II: Beyond Bernoulli and formulation as a MCDP}

\paragraph{Problem Statement}
Same as before, but now a type $j$ is associated with values $r_{j1},r_{j2}>0$ and $q_{j1},q_{j2}\in[0,1]$ such that $q_{j1}+q_{j2}\leq 1$.
The reward of a type $j$ is $r_{jk}$ with probability $q_{jk}$, $k\in\crl{1,2}$, and $0$ with probability $q_{j0}\defeq1-q_{j1}-q_{j2}$.
There is a conceptual gap from this case with three values and the binary (active/inactive), but the extension from three to multiple values is straightforward.
We denote $\bar r_j\defeq r_{j1}q_{j1}+r_{j2}q_{j2}$.

Let us define an \emph{input process} indexed by $t\in[2T]$.
At even times, the input reveals the type $j$ and at odd times it reveals the actual realization given probabilities $q_{jk}$.
Formally, let $(\xit)$ be a Markov chain in the state space $[n]\cup\crl{(j,k):j\in[n],k=0,1,2}$.
The transition probabilities are given by $P_{j,(j,k)}=q_{jk}$, $P_{(j,k),j'}=p_{j'}$.
For $\xi^{2t-1}=(j,k)$ we define $r(\xi^{2t-1})=r_{jk}$.

The state space for our problem can be described as $\St=\crl{(b_h,b_p),(b_h,b_p,\diamond):b_h,b_p\in\N,\diamond\in \crl{\texttt{a,p,r}}}$, see \cref{fig:probing_states}.
The component $\diamond$ indicates if we are in the accept, probe or reject stage.
There is never a $\diamond$ at times $2t$ and, at $2t-1$, feasible controls are different for distinct $\diamond$.

\begin{figure}
\centering
\scalebox{0.9}{%
\input{img/probing_states}
}
\caption{Transitions for the online probing problem. 
Numbers below the arrows represent the reward of a transition.
At time $2t$ the possible actions are: accept, probe and reject.
At $2t-1$, depending on $\diamond$, we can either accept or reject and $b_h$ may be discounted.
}
\label{fig:probing_states}
\end{figure}

The natural relaxation is similar to \cref{eq:probing_relaxation}, but it presents a new challenge.
For binary cases, it was obvious that, if \off wants to probe an arrival, he will always accept it if it is active.
In this case it could be that \off probes $j$ in hope of observing $r_{j2}$ and will reject $r_{j1}$.
We must, therefore, reason about \off's actions at all times, instead of just at even times as before and concluding with structural properties.

Let $\calX\defeq\Rp^{4n}$ be the space for the decision variables.
Note that, by definition, $Z_j(2t-1)=Z_j(2t)-\In{\xi^{2t}=j}$.
In \cref{eq:probing_even,eq:probing_odd} we present two similar functions, observe that they differ by the last two constraints only.

\begin{equation}\label{eq:probing_even}
\begin{array}{rrll}
\varphi(2t,b,Z,\xi^{2t}) \; \max & \multicolumn{3}{l}{\sum_{j,k}r_{jk}(x_{(j,k)a}+q_{jk}x_{ja})} \\
\text{s.t.}& \sum_j x_{jp} & \leq b_p \\
& \sum_{j,k}x_{(j,k)a}+\sum_jx_{ja} &\leq b_h \\
&  x_{ja} +x_{jp} & \leq Z_j(t)  & j\in [n] \\
& x_{(j,k)a} &\leq q_{jk}x_{jp} &  j \in [n],k\in [2].
\end{array}
\end{equation}
\begin{equation}\label{eq:probing_odd}
\begin{array}{rrll}
\varphi(2t-1,b,Z,\xi^{2t-1}) \; \max & \multicolumn{3}{l}{\sum_{j,k}r_{jk}(x_{(j,k)a}+q_{jk}x_{ja})} \\
\text{s.t.}& \sum_j x_{jp} & \leq b_p \\
& \sum_{j,k}x_{(j,k)a}+\sum_jx_{ja} &\leq b_h \\
&  x_{ja} + x_{jp} & \leq Z_j(t)-\In{\xi^{2t}=j}  & j\in [n] \\
& x_{(j,k)a} &\leq q_{jk}x_{jp} + \In{\xi^{2t-1}=(j,k)}&  j \in [n],k\in [2].
\end{array}
\end{equation}

An argument as in \cref{prop:probing_relaxation} shows that $\voff(t,b)\leq \varphi(t,b,Z,\xit)$.
We can pretend, thus, that \off agrees to receive rewards according to $\varphi$.
To make this precise, we need a map between two worlds, since \onl decides on actions \texttt{a,p,r} and \off on points in $\calX$.

\begin{lemma}
Let $\bar X\in\calX$ be a maximizer of $\varphi(2t,b,Z,\xi^{2t})$.
Say $\xi^{2t}=j$, then, at time $2t$, \off is satisfied with the following actions:
\begin{enumerate}
\item Accept if $\bar X_{ja}\geq 1$
\item Probe if $\bar X_{jp}\geq 1$
\item Reject if $\bar X_{ja}+\bar X_{jp}\leq Z_j(t)-1$
\end{enumerate}
At time $2t-1$, \off is satisfied accepting type $(j,k)$ in a probing stage if $\bar X_{jp}\geq 1$ and $\bar X_{(j,k)a}\geq 1$.
\end{lemma}
\begin{proof}
\off is satisfied solving the following problem, where the actions correspond, respectively, to accept, probe and reject
\[
\max\crl{\bar r_j+\E_{\xi^{2t-2}}[\varphi(2t-2,b-e_h,Z,\xi^{2t-2})], \E_{\xi^{2t-2}}[\varphi(2t-1,b-e_p,Z,\xi^{2t-1})], \E_{\xi^{2t-2}}[\varphi(2t-2,b,Z,\xi^{2t-2})]}.
\]
For the accept and reject options the expectation is redundant, since all the necessary information to solve the problem is given in $Z$.
This is not true for the probe option, since \off is unaware of $\xi^{2t-2}$.
By Jensen's inequality, $\E_{\xi^{2t-2}}[\varphi(2t-1,b-e_p,Z,\xi^{2t-1})]\leq \varphi(2t-1,b-e_p,Z,\E[\xi^{2t-1}])$, where we are abusing notation to denote that in \cref{eq:probing_odd} we are replacing $\In{\xi^{2t-1}=(j,k)}$ by $q_{jk}$.
It follows now that \off is also satisfied solving the problem 
\[
\max\crl{\bar r_j+\varphi(2t-2,b-e_h,Z,\xi^{2t-2}), \varphi(2t-1,b-e_p,Z,\E[\xi^{2t-1}])], \varphi(2t-2,b,Z,\xi^{2t-2})}.
\]
Finally, we can relate each of these options with the optimum of $\varphi(2t,b,Z,\xi^{2t})$ in the cases (1), (2) and (3).
The argument at odd times $2t-1$ is similar. 
\end{proof}

>>>>

\section{Stochastic Knapsack With Bayesian Learning}

\paragraph{Problem Statement}
Items with known reward $r$ and random weight $W$ can be placed in a knapsack of capacity $B$.
At every time $t$, an arrival $j\in[n]$ is drawn with some known distribution, which generates a known reward of $r_j$.
The weight $W_j$ is a random variable, which is revealed only after the decision of accept/reject has been made.
An arrival can be accepted only if the weight is a.s.\ smaller than the remaining capacity.
Furthermore, the distribution of $W_j$ depends on an unknown parameter $\theta_j$ in the parametric form $W_k|\theta_j\sim F_{\theta_j}$.
Before the process starts, we are given a prior distribution $\Theta_j$ for the parameter of each type.

We assume that, at the end of each period, we observe the realization of both accepted and rejected items.
Let $\calF_t$ be the $\sigma$-algebra generated by the process, which includes the observations of weights up to, and including, time $t$.
We define $\calF_{T+1}\defeq\varnothing$.
We assume the updated average weight $\hat W_{jt} \defeq  \E[W_j|\calF_{t+1}]$ can be computed efficiently.

The intuition behind our relaxation is as follows.
At any time $t$, we give more power to the decision maker by (1) revealing the future arrivals $Z(t)$ and (2) replacing the items by deterministic copies having weight equal to $\hat W_{tj}$, instead of random.
The relaxation is presented in \cref{eq:knapsack_lp}.
Observe that the dependence in the arrival $\xit$ is ``absorbed'' in $Z(t)$.

\begin{equation}\label{eq:knapsack_lp}
\begin{array}{rrll}
\varphi(t,b,\xit) \; = \; \max & \multicolumn{3}{l}{\sum_{j}r_{j}x_j} \\
\text{s.t.}& \sum_{j}\hat W_{jt}x_j  &\leq b  \\
&  x_j & \leq Z_j(t)  & j\in [n] \\
& x&\geq 0.
\end{array}
\end{equation}
